Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-11-18 bbbq_article.tex
All Maksim's feedback is in the document, progress #253

Epitope prediction is important to understand the immune system 
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism. Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes. Therefor, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms. Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

Although our data show that 
TMH-derived epitopes are presented in all classical MHC-I and MHC-II haplotypes, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown \cite{bianchi2017}. 
Most prominently, the fundamental principles of 
how TMHs are extracted from their hydrophobic lipid environments 
into the aqueous vacuolar lumen, 
and their prior or subsequent proteolytic processing are unresolved.

@@

In general, one might expect that evolutionary selection results in
an immune system that is most attentive for protein regions that are
essential for the survival, proliferation and/or virulence or pathogenic microbes, 
as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that change its 
virulence \cite{velazquez2020positive}.
These regions, however, may only account for a small part of a pathogen's proteome.
Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible 
to provide innate immune responses against such rare essential protein regions, 
as suggested in a study on influenza \cite{han2019individual},
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

@@

Evolutionary selection of pathogens by a host's immune system,
however, is likelier to occur for protein patterns that are general,
over patterns that are rare.
While essential catalytic sites in a pathogenic proteome
might be relatively rare, TMHs are common and thus might be a more feasible 
target for evolution to respond to.
Indeed, we have found the signature of evolution when both factors,
that is, TMHs and catalytic sites are likely to co-occur,
which is in TMPs that span the membrane at least twice.
In contrast to single-spanners, where we found no significant evolutionary conservation, the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions. Likely, the TMHs in many multi-spanners need to interact which each other for correct protein structure and function and they might hence be more structurally constrained compared to the TMHs of single-spanners.
Thus, we speculate that the human immune system is more attentive 
towards TMHs in multi-spanners, as these are evolutionarily more conserved.

>>>>

Epitope prediction is important to understand the immune system function
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism. Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes. Therefore, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms. Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

Although our data show that 
TMH-derived epitopes are presented in all classical MHC-I and MHC-II haplotypes, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown \cite{bianchi2017}. 
Most prominently, the fundamental principles of 
how TMHs are extracted from their hydrophobic lipid environments 
into the aqueous vacuolar lumen, 
leading to subsequent proteolytic processing are unresolved.

@@

In general, one might expect that evolutionary selection shapes an immune system 
where surveillance is directed towards protein regions 
essential for the survival, proliferation and/or virulence or pathogenic microbes, 
as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is directed upon residues that change its 
virulence \cite{velazquez2020positive}.
These regions, however, may only account for a small part of a pathogen's proteome.
Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible 
to provide innate immune responses against such rare essential protein regions, 
as suggested in a study on influenza \cite{han2019individual},
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

@@

Evolutionary selection of pathogens by a host's immune system,
however, is more likely to occur for protein patterns that are general,
over patterns that are rare.
While essential catalytic sites in a pathogenic proteome
might be relatively rare, TMHs are common and thus might be a more feasible 
target for evolution to respond to.
Indeed, we have found the signature of evolution when both factors,
that is, TMHs and catalytic sites are likely to co-occur,
which is in TMPs that span the membrane at least twice.
In contrast to single-spanners, where we found no significant evolutionary conservation, the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions. Likely, the TMHs in many multi-spanners need to interact which each other for correct protein structure and function and they might hence be more structurally constrained compared to the TMHs of single-spanners.
Thus, we speculate that the human immune system is more attentive 
towards TMHs in multi-spanners, as these are evolutionarily more conserved.
